---
title: "Solution to Exercise 1 & 2"
author: "Christopher Maronga - R Programmer"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: vignette
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      message = F,
                      warning = F,
                      eval = F)
```

## Introduction

Welcome to part one of the exercises, connecting to `RDBMS`. This is short exercise with the following objectives:-

- Connecting to a `MySQL` database using `odbc` driver.
- Connecting to a `MySQL` database using a `DBI` back-end package
- Exploring database contents using the established connection.

We have set-up 3 `MySQL` databases for use in this exercise and subsequent exercises throughout the workshop. Below are the names and credentials of the databases to connect to

(i) Database name :`nhsr_conference`
      - **host/server** : `nhsr-conference.cjmn9gy7nffs.us-east-1.rds.amazonaws.com`
      - **user/Uid** : `user`
      - **Password** : `user`

(ii) Database name :`nycflights13`
      - **host/server** : `nhsr-conference.cjmn9gy7nffs.us-east-1.rds.amazonaws.com`
      - **user/Uid** : `user`
      - **Password** : `user`

(ii) Database name :`SUCE`
      - **host/server** : `109.74.194.173`
      - **user/Uid** : `opendata`
      - **Password** : `letmein`

**NOTE:** In connecting to databases, sometimes it's necessary to specify a `port`, if that's the case during this exercises, kindly specify `port = 3306`

  
## Load packages

Load packages you need to connect, extract and wrangle data from your databases

```{r}
# loading packages
```




## `r kableExtra::text_spec("Question one - Connecting to a database", color = "steelblue")`


### Using an `odbc` driver



### Using a `DBI` backend package


### Exploring using the connections




##  `r kableExtra::text_spec("Question two - Extracting data from a database", color = "steelblue")`



### Using `DBI` functions



### Using `dplyr` syntax



### Using a `SQL` chunck





























