---
title: "Solution to Exercise 1 & 2"
author: "Christopher Maronga - R Programmer"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: vignette
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      message = F,
                      warning = F,
                      eval = F)
```

## Introduction

Welcome to part one of the exercises, connecting to `RDBMS`. This is short exercise with the following objectives:-

- Connecting to a `MySQL` database using `odbc` driver.
- Connecting to a `MySQL` database using a `DBI` back-end package
- Exploring database contents using the established connection.

We have set-up 3 `MySQL` databases for use in this exercise and subsequent exercises throughout the workshop. Below are the names and credentials of the databases to connect to

(i) Database name :`nhsr_conference`
      - **host/server** : `nhsr-conference.cjmn9gy7nffs.us-east-1.rds.amazonaws.com`
      - **user/Uid** : `user`
      - **Password** : `user`

(ii) Database name :`nycflights13`
      - **host/server** : `nhsr-conference.cjmn9gy7nffs.us-east-1.rds.amazonaws.com`
      - **user/Uid** : `user`
      - **Password** : `user`

(ii) Database name :`SUCE`
      - **host/server** : `109.74.194.173`
      - **user/Uid** : `opendata`
      - **Password** : `letmein`

**NOTE:** In connecting to databases, sometimes it's necessary to specify a `port`, if that's the case during this exercises, kindly specify `port = 3306`

  
## Load packages

Load packages you need to connect, extract and wrangle data from your databases

```{r}
# load packages
library(DBI)        # Main database interface for R
library(dbplyr)     # Handles dplyr translation to SQL
library(tidyverse)  # tidyverse world of package (including dplyr, ggplot2, etc.)
library(RSQLite)    # DBI back-end package for connecting to SQLite
library(odbc)       # DBI back-end  to connect to most RDBMS & SQL server
library(RMySQL)     # DBI back-end package for connecting to MySQL
library(redcapAPI)  # Connecting to REDCap database
```




## `r kableExtra::text_spec("Question one - Connecting to a database", color = "steelblue")`


### Using an `odbc` driver

```{r}
# Task : Connect to `nycflights13` MySQL database;

# (i) Use an `odbc` driver, name the connection `nyc_odbc.conn`
# Hint: 
sort(unique(odbcListDrivers()[[1]])) # Run this to see available odbc drivers


# establishing a database connection
nyc_odbc.conn <- dbConnect(odbc(),                                       # Open a database connection
                             Driver = "MySQL ODBC 8.0 Unicode Driver",   # Specify driver name (), for 
                             Server = "nhsr-conference.cjmn9gy7nffs.us-east-1.rds.amazonaws.com",   # Server address
                             Database = "nycflights13",               # Database name
                             Uid = "user",                            # User identification
                             Pwd = "user" )


```



### Using a `DBI` backend package

```{r}
# (ii) Use a `DBI` backend package for MySQL, name the connection `nyc_dbi.conn`

nyc_dbi.conn <- dbConnect(MySQL(),  # DBI compliant driver for MySQL db
                       host = "nhsr-conference.cjmn9gy7nffs.us-east-1.rds.amazonaws.com",     # host name/server address
                       dbname = "nycflights13",   # database name
                       user = "user",     # user name
                       password = "user") # password
```


### Exploring using the connections

```{r}
# (iii) Explore the connection(s)

# How many tables are in the connection `nyc_odbc.conn`?
dbListTables(conn = nyc_odbc.conn)


# How many tables are in the connection `nyc_dbi.conn`?
dbListTables(conn = nyc_dbi.conn)

## Using either of the database connections above, how many variables are in table airports table?(tbl_airports)
dbListFields(conn = nyc_odbc.conn, "tbl_airports")

```



##  `r kableExtra::text_spec("Question two - Extracting data from a database", color = "steelblue")`



### Using `DBI` functions

Use the combination of `dbSendQuery()` to execute a SQL query in the database and return some results. Followed by the function `dbFetch()` to fetch the results returned in step 1 into actual data in memory

```{r}
# establishing a connection with the database
nhsr_db.conn <- dbConnect(MySQL(),  
                       host = "nhsr-conference.cjmn9gy7nffs.us-east-1.rds.amazonaws.com",
                       dbname = "nhsr_conference",   
                       user = "user",   
                       password = "user") 

# list tables in the databse
dbListTables(nhsr_db.conn)


# Step 1
qry_results <- dbSendQuery(conn = nhsr_db.conn,
            statement = "SELECT * FROM tbl_clinical")

# Step 2
clinical_opt1 <- dbFetch(res = qry_results)

```


```{r}
# (b) use dbGetQuery() to execute the SQL above (i) and assign to `clinical_opt2`
## NOTE: this is equivalent to using the function dbReadTable()

clinical_opt2 <- dbGetQuery(conn = nhsr_db.conn,
                            statement = "SELECT * FROM tbl_clinical")

# also
# clinical_opt2 <- dbReadTable(conn = nhsr_db.conn, 
#                              "tbl_clinical")  # specify table name

```


### Using `dplyr` syntax

```{r}
# use the tbl() function and collect() to complete the set of exercise below

# (i) Extract the lab data and assign it to `lab_data` in your enviroment

# remind yourself how many tables in the database
dbListTables(conn = nhsr_db.conn)

# extract lab data
lab_data <- tbl(nhsr_db.conn, "tbl_lab") %>% 
  collect() # collect() pulls data into memory


# (ii) Extract patients whose sex == 1 and site == 3, assign to `site1_male.df`

site1_male.df <- tbl(nhsr_db.conn, "tbl_baseline") %>% 
  filter(sex == 1, site == 3) %>% 
  collect()


# (iii) Use an appropriate join to extract and combine all tables of the database, assign to `study_raw`
study_raw <- tbl(nhsr_db.conn, "tbl_baseline") %>% 
  left_join(
    tbl(nhsr_db.conn, "tbl_clinical"),
    by = "subjid"
  ) %>% 
  left_join(
    tbl(nhsr_db.conn, "tbl_lab"),
    by = "subjid"
  ) %>% 
  collect()

```



Using a connection created by any method, you can write a bar SQL code in sql_code chunk as shown below

### Using a `SQL` chunck

```{sql, connection = nhsr_db.conn, output.var = "study_raw_opt2"}

-- This is bare SQL code and the results is saved in "study_raw_opt2" to be used 
SELECT * 
FROM tbl_baseline
LEFT JOIN tbl_clinical ON tbl_baseline.subjid = tbl_clinical.subjid
LEFT JOIN tbl_lab ON tbl_baseline.subjid = tbl_lab.subjid
```

You can also use what we call table alias

```{sql, connection = nhsr_db.conn, output.var = "study_raw_opt2"}

-- This is bare SQL code and the results is saved in "study_raw_opt2" to be used 
SELECT * 
FROM tbl_baseline AS base
LEFT JOIN tbl_clinical AS cl ON base.subjid = cl.subjid
LEFT JOIN tbl_lab AS lb ON base.subjid = lb.subjid
```





























