---
title: "Solution to Exercise 3"
author: "Christopher Maronga - R Programmer"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: vignette
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      message = F,
                      warning = F,
                      eval = F)
```

## Introduction

Welcome to part two of the exercises, connecting to `REDCap`. Objectives of the exercise:-

- Connecting `R` and `REDCap` using an `API`
- Exporting data from `REDCap`
      - `All` records at a go.
      - Records by `event`
      - Records by `instrument/form`

We have set up `one` REDCap database to use in this exercise. Below are the details to connect to the database.

(i) Database link :`https://uat.chainnetwork.org/`
      - **url** : `https://uat.chainnetwork.org/redcap/api/`
      - **API token** : `125AAA97191E444C80E5A796C40F1E52`


**NOTE:** You only need two parametres to connect to a REDCap database, the `url link `and the API `token/code`

  
## Load packages

Load packages you need to connect, extract and wrangle data from your databases

```{r}
# loading packages
library(DBI)        # Main database interface for R
library(dbplyr)     # Handles dplyr translation to SQL
library(tidyverse)  # tidyverse world of package (including dplyr, ggplot2, etc.)
library(RSQLite)    # DBI back-end package for connecting to SQLite
library(odbc)       # DBI back-end  to connect to most RDBMS & SQL server
library(RMySQL)     # DBI back-end package for connecting to MySQL
library(redcapAPI)  # Connecting to REDCap database
```




## `r kableExtra::text_spec("Question one - Create a connection", color = "steelblue")`

```{r}
# create/establish a connection
redcap_conn <- connection_name <- redcapConnection(url = "https://uat.chainnetwork.org/redcap/api/", # url
                               token = "125AAA97191E444C80E5A796C40F1E52")  # token
```

### Exploring the database

You can use the connection to explore the databse further

```{r}
# how many events and names?
events <- exportEvents(redcap_conn) # these are just timepoints

# how many instruments and names?
instruments <- exportInstruments(redcap_conn)

# users and privileges?


# metadata?

```


##  `r kableExtra::text_spec("Question two - Export data", color = "steelblue")`


### All data at a go

```{r}
# This approach for a small study, huge data will time-out connection

clinical_data <- exportRecords(rcon = redcap_conn, 
                               factors = TRUE)

# Can you sport the difference?
clinical_data <- exportRecords(rcon = redcap_conn, 
                               factors = FALSE)
```


### Export data by event

```{r}
# remind yourself on events
events

# export baseline event
baseline <- exportRecords(rcon = redcap_conn, factors = TRUE, 
                          fields = NULL,  forms = NULL, records = NULL, 
                          events = "baseline_arm_1",
                          labels = TRUE, dates = TRUE, 
                          survey = TRUE, dag = TRUE,  
                          checkboxLabels = FALSE)
# clean columns
baseline <-  baseline[, colSums(is.na(baseline)) < nrow(baseline)] ## exclude NA filled columns


# export month 3 data
month_3 <- exportRecords(rcon = redcap_conn, factors = TRUE, 
                          fields = NULL,  forms = NULL, records = NULL, 
                          events = "month_3_arm_1",
                          labels = TRUE, dates = TRUE, 
                          survey = TRUE, dag = TRUE,  
                          checkboxLabels = FALSE)
# clean columns
month_3 <-  month_3[, colSums(is.na(month_3)) < nrow(month_3)] ## exclude NA filled columns

# export month 6 data
month_6 <- exportRecords(rcon = redcap_conn, factors = TRUE, 
                          fields = NULL,  forms = NULL, records = NULL, 
                          events = "month_6_arm_1",
                          labels = TRUE, dates = TRUE, 
                          survey = TRUE, dag = TRUE,  
                          checkboxLabels = FALSE)
# clean columns
month_6 <-  month_6[, colSums(is.na(month_6)) < nrow(month_6)] ## exclude NA filled columns

# export study exit
study_exit <- exportRecords(rcon = redcap_conn, factors = TRUE, 
                          fields = NULL,  forms = NULL, records = NULL, 
                          events = "study_end_arm_1",
                          labels = TRUE, dates = TRUE, 
                          survey = TRUE, dag = TRUE,  
                          checkboxLabels = FALSE)
# clean columns
study_exit <-  study_exit[, colSums(is.na(study_exit)) < nrow(study_exit)] ## exclude NA filled columns

# merge/join by subjid
redcap_clinical_data <- reduce(
  list(baseline,
       month_3,
       month_6,
       study_exit),
  left_join,
  by = "record_id"
)

```


### Export data by instrument/form

```{r}
# Try this at your free time and check how it compares with the previous methods.

```


### Have fun playing around with the database

```{r}
# Have fun trying out new combination of extracting instruments and events

```


























