---
title: "Solution to Exercise 3"
author: "Christopher Maronga - R Programmer"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: vignette
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      message = F,
                      warning = F,
                      eval = F)
```

## Introduction

Welcome to part two of the exercises, connecting to `REDCap`. Objectives of the exercise:-

- Connecting `R` and `REDCap` using an `API`
- Exporting data from `REDCap`
      - `All` records at a go.
      - Records by `event`
      - Records by `instrument/form`

We have set up `one` REDCap database to use in this exercise. Below are the details to connect to the database.

(i) Database link :`https://uat.chainnetwork.org/`
      - **url** : `https://uat.chainnetwork.org/redcap/api/`
      - **API token** : `125AAA97191E444C80E5A796C40F1E52`


**NOTE:** You only need two parametres to connect to a REDCap database, the `url link `and the API `token/code`

  
## Load packages

Load packages you need to connect, extract and wrangle data from your databases

```{r}
# loading packages
```




## `r kableExtra::text_spec("Question one - Create a connection", color = "steelblue")`


### Exploring the database




##  `r kableExtra::text_spec("Question two - Export data", color = "steelblue")`



### All data at a go



### Export data by event



### Export data by instrument/form





























