---
title: <center><strong> Working with databases in R - Prerequisites </strong></center>
author: <center><strong> Christopher Maronga </strong></center>
date: <center><strong> `r format(Sys.Date(), '%A, %d %B %Y')` </strong></center>
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      message = F,
                      warning = F,
                      eval = F)
```

## Introduction

Thank you for planning to attend the upcoming NHS-R Conference and choosing Working with Databases in R workshop. To get started and make sure you are able to follow along during the live coding, we would advise you to install in advance the following list of packages that we will be working with.

- `tidyverse`
- `DBI` 
- `odbc` 
- `RMySQL`
- `RSQLite`
- `REDCapR`
- `redcapAPI`

You can install a package by running following code in your R console, `install.packages("package_name")`. For instance, if you want to install `tidyverse`, you could just run `install.packages("tidyverse")`.

You could also copy, paste and run the code below to install and load all the above packages.

\vspace{5mm}


```{r packages}
# checks if pacman is intalled
if (!require(pacman)) install.packages("pacman")

# manages the rest of the packages (installs and loads)
pacman::p_load(tidyverse, DBI, odbc, 
               RMySQL, RSQLite, REDCapR, 
               redcapAPI)

```


## Databases

We have prepared 3 databases for use during the meetup. These will be utilized during the exercises and feel free to use them while practicing code on your own time.

You can download the folder containing these databases [here](https://www.dropbox.com/sh/g04mokleaca8xsk/AADZgkiS2xJZdAfzD6DWATiza?dl=0).

- `naidemostudy` - Self created imaginary longitudinal study
- `nyc_flights` - Dataset from `nycflights13` package. You can find more information [here](https://github.com/tidyverse/nycflights13)
- `world` - Demo dataset that comes with SQL installation



**See you soon!**
















